### 2021-12-11 documents/sections/extensions.tex
extension finetuning

\section{Inflections and verbs: innovation and resistance}
\label{sec:data}
As was shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
\cref{sec:extensions} presents the six known incomplete extensions, the prefixes they introduced and the verbs they spared.
Since the latter show considerable etymological overlap across languages, the resistant verbs are compared and reconstructed in \cref{sec:verbs}.
In addition, their reflexes which did get affected by one of the extensions are identified, where existent.
%The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.

\subsection{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
\label{sec:extensions}
As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
Of these extensions, half can be reconstructed to intermediate proto-languages, while the others happened in pre-modern stages of single languages.
The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary:
\cref{sec:pekodian} details the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
\cref{sec:waiwaian} treats the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
\cref{sec:taranoan} focuses on innovative \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers only found in single languages:
\obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.

\subsubsection{\PPek \rc{k-}}
\label{sec:pekodian}
%The Pekodian branch was suggested by \textcite{meira2005southern}, as the result of fieldwork on \bakairi by Meira and the availability of more material on \ikpeng. % \parencites{pacheco1997relativizacao}{ikpengpacheco1997}{campetela1997analise}{pacheco1998nominalizacao}{ikpengpacheco2001}{campetela2002aspectos}{pacheco2002aspectos}{pacheco2003intransitivos}.
The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
The contribution establishing the branch \parencite{meira2005southern} focuses on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.

@@

In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{
	It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
	Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
	If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
}
The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}.

\ex<bak-33>\bakairi \parencite[][4]{meira2003bakairi}\\
\begingl
\gla w-i-də//
\glb \gl{1}\gl{s_a_}-bathe-\gl{imm}//
\glft \qu{I bathed}//
\endgl
\xe
%
%Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
Inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.

@@

All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}.
The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
While reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{to be} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.

@@

\ex<ara-123>\arara \parencite[][150]{alves2017arara}\\
\begingl
\gla m-odebɨ-nɨ//
\glb \gl{2}\gl{s_a_}-come-\gl{rec}//
\glft \qu{You came.}//
\endgl
\xe
%
In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara should be reconstructed to \PPek, while \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.

\subsubsection{\PWai \rc{k-}}
\label{sec:waiwaian}
This extension, one of the Parukotoan innovations discussed in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
In \PWai, the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar was further extended to \gl{1}\gl{s_a_}.
For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
\input{floats/pwaireg}

@@

There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
The two forms of \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}
%
%\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
%\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
%\a \obj{o-\emp{w}to-t͡ʃhe} \qu{after I went} \parencite[165]{waiwaihawkins1998}
%\a \obj{kɨ-\emp{w}to-me} \qu{for us to go} \parencite[204]{waiwaihawkins1998}
%\xe
%
In any case, the first person form \hixka \qu{to go} clearly points to \PWai \rc{wɨ-tom-}.

@@

\subsubsection{\PTir \rc{t-}}
\label{sec:taranoan}
The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, with \carijo as a more distant member.
\textcite{meira1998proto} contributes an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
\input{floats/ptirreg}
The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
In fact, even the issue of \emph{when} this happened is open.
It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
Here, a conservative stance is taken and the innovation is arbitrarily assumed to be \PTir.
This decision does not affect the results of this study.}

@@

Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
To this list the second copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is plausibly identified as a reflex of \rc{wɨ-} by 
\textcite[113]{meira1998proto}, which is supported by additional evidence:
Both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.

\pex<exp>\akuriyo
\a<aku-164>
\begingl
\gla wi-toka//
\glb \gl{1}>\gl{3}-hit//
\glft \qu{I hit him/her.} \parencite[][86]{gildea1994akuriyo}//
\endgl
\a<aku-159>
\begingl
\gla kəʔ-eepɨ//
\glb \gl{1+2}-come//
\glft \qu{We came.} \parencite[][114]{meira1998proto}//
\endgl
\xe
%
%For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
%The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
%We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
%Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
%This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
%
%
%

For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).

>>>>

\section{Inflections and verbs: innovation and resistance}
\label{sec:data}
As shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
\cref{sec:extensions} presents the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
Since the latter show considerable etymological overlap across languages, these resistant verbs are compared and reconstructed in \cref{sec:verbs}.
Where existent, their reflexes which did get affected by one of the extensions are identified.
%The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.

\subsection{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
\label{sec:extensions}
As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets have in common that they introduced innovative first person markers on \gl{s_a_} verbs.
Of these extensions, half can be reconstructed to intermediate proto-languages, and half happened in pre-modern stages of single languages.
The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary.
\cref{sec:pekodian} details the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
\cref{sec:waiwaian} treats the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
\cref{sec:taranoan} focuses on innovative \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers only found in single languages:
\obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.

\subsubsection{\PPek \rc{k-}}
\label{sec:pekodian}
%The Pekodian branch was suggested by \textcite{meira2005southern}, as the result of fieldwork on \bakairi by Meira and the availability of more material on \ikpeng. % \parencites{pacheco1997relativizacao}{ikpengpacheco1997}{campetela1997analise}{pacheco1998nominalizacao}{ikpengpacheco2001}{campetela2002aspectos}{pacheco2002aspectos}{pacheco2003intransitivos}.
The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
The contribution establishing the branch \parencite{meira2005southern} focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.

@@

In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
	Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
	If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
}
The first group is illustrated with \obj{i} \qu{to bathe} \exref{bak-33}.

\ex<bak-33>\bakairi \parencite[][4]{meira2003bakairi}\\
\begingl
\gla w-i-də//
\glb \gl{1}\gl{s_a_}-bathe-\gl{imm}//
\glft \qu{I bathed}//
\endgl
\xe
%
%Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
However, inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.

@@

All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.
There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
While reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{to be} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.

@@

\ex<ara-123>\arara \parencite[][150]{alves2017arara}\\
\begingl
\gla m-odebɨ-nɨ//
\glb \gl{2}\gl{s_a_}-come-\gl{rec}//
\glft \qu{You came.}//
\endgl
\xe
%
In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in Taranoan (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara can be reconstructed to \PPek, meaning that \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.

\subsubsection{\PWai \rc{k-}}
\label{sec:waiwaian}
This extension, one of the Parukotoan innovations shown in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
\PWai further extended the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar to \gl{1}\gl{s_a_}.
For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
\input{floats/pwaireg}

@@

There are three more verbs which did not take innovative \rc{k-} in \PWai \pcref{tab:pwaiverbs}.
The two forms for \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}
%
%\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
%\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
%\a \obj{o-\emp{w}to-t͡ʃhe} \qu{after I went} \parencite[165]{waiwaihawkins1998}
%\a \obj{kɨ-\emp{w}to-me} \qu{for us to go} \parencite[204]{waiwaihawkins1998}
%\xe
%
In any case, the first person form \hixka \qu{to go} clearly points to \PWai \rc{wɨ-tom-}.

@@

\subsubsection{\PTir \rc{t-}}
\label{sec:taranoan}
The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a more distant member.
\textcite{meira1998proto} contributes an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
\input{floats/ptirreg}
The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
In fact, even the issue of \emph{when} this happened is open.
It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
Here, a conservative stance is taken and the innovation is arbitrarily assumed to be \PTir.
This decision does not affect the results of this study.}

@@

Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
The other copular root \rc{eʔi} can be added as a fifth verb, whose \trio reflex retains first person \obj{w-}.
The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
\textcite[113]{meira1998proto}, which is supported the fact that both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.

\pex<exp>\akuriyo
\begin{multicols}{2}
\a<aku-164>
\begingl
\gla wi-toka//
\glb \gl{1}>\gl{3}-hit//
\glft \qu{I hit him/her.} \parencite[][86]{gildea1994akuriyo}//
\endgl
\a<aku-159>
\begingl
\gla kəʔ-eepɨ//
\glb \gl{1+2}-come//
\glft \qu{We came.} \parencite[][114]{meira1998proto}//
\endgl
\end{multicols}
\xe
%
%For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
%The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
%We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
%Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
%This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
%
%
%

For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on such a paradigmatic pattern in \trio and the vowel length in \akuriyo.
Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-jəpɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
